---
title: "Deskriptivní statistiky časových řad"
author: "Daniel Bartušek"
output: html_document
runtime: shiny
---
```{r setting, echo = F}
rm(list = ls())
options(scipen = 100, digits = 8)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(echo = TRUE)
quiet <- function(x) { 
  sink(tempfile()) 
  on.exit(sink()) 
  invisible(force(x)) 
}
```

```{r functions,echo=F}
format_cells <- function(df, rows ,cols, value = c("italics", "bold", "strikethrough")){
  map <- setNames(c("*", "**", "~~"), c("italics", "bold", "strikethrough"))
  markup <- map[value]
  for (r in rows){
    for(c in cols){
      df[[c]] <- as.character( df[[c]])
      df[r, c] <- paste0(markup, df[r, c], markup)
    }
  }
  return(df)
}

best_garch<-function(fit){
  p_max <- 5
  q_max <- 5
  aic_min <- Inf
  best_p <- 0
  best_q <- 0
  x <- fit$x %>%as.vector()
  ar <- ifelse(length(fit$coef)>0,fit$arma[2],0)
  ma<- ifelse(length(fit$coef)>0,fit$arma[1],0)
  for (i1 in 1:p_max) {
    for (i2 in 1:q_max) {
      model_specification <- ugarchspec(mean.model = list(armaOrder = c(ar,ma), include.mean = TRUE),
                                        variance.model = list(garchOrder = c(i1, i2)))
      fit <- ugarchfit(spec = model_specification, data = x)
      inf_crit <- infocriteria(fit)[1]
      aic_min <- ifelse(inf_crit < aic_min, inf_crit, aic_min)
      best_p <- ifelse(inf_crit == aic_min, i1, best_p)
      best_q <- ifelse(inf_crit == aic_min, i2, best_q)
    }
  }

  return(c(best_p,best_q))
}
```

```{r packages, echo = F }
library(ggplot2)
library(readxl)
library(ggpubr)
library(kableExtra)
library(stargazer)
library(moments)
library(tseries)
library(lmtest)
library(stringr)
library(frequencyConnectedness)
library(imputeTS)
library(ggcorrplot)
library(lubridate)
library(dplyr)
library(reshape2)
library(strucchange)
library(parallel)
library(BigVAR)
library(scales)
library(here)
library(gridExtra)
library(rmgarch)
library(aTSA)
library(forecast)
library(FinTS)
```

```{r date_ranges,echo=F}
gfc_from<-"2007-06-29"
post_gfc_from<-"2009-02-06"
low_oil_from<-"2014-11-30"
fed_rates_from<-"2017-08-10"
covid_from<-"2020-01-01" 
gfc_to<-"2009-02-05"
post_gfc_to<-"2014-11-29"
low_oil_to<-"2017-08-09"
fed_rates_to<-"2019-12-31"
covid_to<-"2022-01-01"
ranges<-data.frame(cbind(c("GFC", "Post-GFC", "Low oil prices","Fed Rates Increase","Covid"),
                         c(gfc_from,post_gfc_from,low_oil_from,fed_rates_from,covid_from),
                         c(gfc_to,post_gfc_to,low_oil_to,fed_rates_to,covid_to)))
colnames(ranges)<-c("labels","from","to")
```


```{r load_data, echo=F}
pca_corr_norm<-readRDS("data/pca/pca_corr.rds")
pca_corr_ret<-readRDS("data/pca/cor_ret_df.rds")
pca_vals<-readRDS("data/pca/pca_vals.rds")
log_ret_df<-readRDS("data/log_ret.rds")
df<-readRDS("data/df.rds")
adcc_fits<-readRDS("data/adcc_fits.rds")
long_log_ret <- melt(log_ret_df, id.vars = "Date")
long_df <- reshape2::melt(df, id.vars = c("Date","Weekday","Holiday"))
stylized_names <-c("Natural Gas","Crude Oil WTI","NY Gasoline","Heating Oil","Diesel")
```

## Prices
```{r prices_stats,echo=F}
  selectInput("period", "Choose a period:",
      list(`Period` = list( "GFC", "Post-GFC", "Low oil prices","Fed Rates Increase","Covid"))
    )

renderPlot({
dt_from<-ranges %>% filter(labels== input$period) %>% select(2) %>% pull()
dt_to<-ranges %>% filter(labels == input$period) %>% select(3) %>% pull()
long_df_aux<-long_df[long_df$Date>=dt_from,]
long_df_aux<-long_df_aux[long_df_aux$Date<=dt_to,]

ggplot(long_df_aux, aes(x = Date, y = value)) +
  geom_line() +
  facet_wrap(~variable, scales = "free") +
  ggtitle("Prices") 
})

renderPlot({
dt_from<-ranges %>% filter(labels== input$period) %>% select(2) %>% pull()
dt_to<-ranges %>% filter(labels == input$period) %>% select(3) %>% pull()
long_df_aux<-long_df[long_df$Date>dt_from,]
long_df_aux<-long_df_aux[long_df_aux$Date<dt_to,]

ggplot(long_df_aux, aes(factor(variable), value)) + 
  geom_boxplot() + 
  facet_wrap(~variable, scale="free")
})

```


```{r bp_comparison, echo= F,eval=F}
breaks<-readRDS("data/struc_change/break_dates.rds")

bp_6<-data.frame(matrix(ncol=9,nrow=5))
bp_7<-data.frame(matrix(ncol=9,nrow=5))
bp_8<-data.frame(matrix(ncol=9,nrow=5))
bp_9<-data.frame(matrix(ncol=9,nrow=5))
rownames(bp_6)<-names(breaks)
rownames(bp_7)<-names(breaks)
for (i in 1:length(breaks)){
  bp_6[i,]<-breaks[[i]][6,]
  bp_7[i,]<-breaks[[i]][7,]
  bp_8[i,]<-breaks[[i]][8,]
  bp_9[i,]<-breaks[[i]][9,]
}

make_bp_df<-function(bp_num){
  bp_name<-data.frame(matrix(ncol=9,nrow=5)) 
  rownames(bp_name)<-names(breaks)
  colnames(bp_name)<-paste("BP",rep(1:length(bp_name)))
  for (i in 1:length(breaks)){
  bp_name[i,]<-breaks[[i]][bp_num,]
  }
  bp_name<-bp_name %>% mutate_all(as.Date)
  bp_sum<-bp_name %>% summarize_all(mean.Date,na.rm = T)
  rownames(bp_sum)<-"Mean"
  bp_name<-rbind(bp_name,bp_sum) %>% format_cells(6, 1:9, "bold") 
  return(bp_name)
}

make_bp_df(6)%>% kbl() %>% kable_classic_2()
make_bp_df(7)%>% kbl() %>% kable_classic_2()
make_bp_df(8)%>% kbl() %>% kable_classic_2()
make_bp_df(9)%>% kbl() %>% kable_classic_2()
```



```{r breakpoints, echo = F,evql=F}
dates<-log_ret_df %>% dplyr::select(1)%>%pull()
char_dates<-as.character(dates)
dates_df<-df %>% dplyr::select(1)%>%pull()
char_dates_df<-as.character(dates_df)
chows<-readRDS("data/struc_change/chow.rds")
breaks<-readRDS("data/struc_change/break_dates.rds")
bp_plots<-readRDS("data/struc_change/bp_plots.rds")
breaks_v2<-readRDS("data/struc_change/break_dates_v2.rds")
bp_plots_v2<-readRDS("data/struc_change/bp_plots_v2.rds")
main_list_plt<-list()
for(i in 4:8){
  main_list_plt<-c(main_list_plt,list(df[i]))
}
names(main_list_plt)<-names(chows)


selectInput("opt_bp", "Choose how many breakpoints to plot:",
      list(`Optimal Breakpoint Number`  = c("Optimal",seq(1,9)))
    )
selectInput("struct_change", "Choose a series:",
      list(`Log Returns Period` = colnames(log_ret_df[-1]))
    )
renderPrint({
   chow<-chows[[input$struct_change]]
    print("Chow test (H0: there are no structural brakes)")
    print(chow)
})
# renderTable({
#    bp_df<-breaks[[input$struct_change]]
#    rownames(bp_df)<-paste(rep(1:nrow(bp_df)),"Breakpoints modeled")
#    colnames(bp_df)<-paste("Breakpoint",rep(1:nrow(bp_df)))
#    bp_df
# },rownames = TRUE)

renderPlot({
  bp_df<-breaks[[input$struct_change]]
  bp_object<-bp_plots[[input$struct_change]]
   index_bp<-which.min(bp_object$RSS[2,])-1
  if(input$opt_bp != "Optimal"){
    index_bp<-as.numeric(input$opt_bp)
  }
  bps<-bp_df[index_bp,] %>% as.character()
  bp_indexes<-match(bps,char_dates_df)
  bp_indexes<-bp_indexes[!is.na(bp_indexes)]
  
  plot.ts(df[[input$struct_change]],xaxt='n')
  axis(1, at=bp_indexes, labels=char_dates_df[bp_indexes])
  abline(v=bp_indexes, col="red")
})

renderPlot({
   bp<-bp_plots[[input$struct_change]]
   plot(bp)
})
```

## Log Returns


```{r corplot_logret,echo=F,message=T,warning=T}
selectInput("period_corrplot", "Choose a period:",
      list(`Log Returns Period` = list("GFC", "Post-GFC", "Low oil prices","Fed Rates Increase","Covid"))
    )

renderPlot({
dt_from<-ranges %>% filter(labels== input$period_corrplot) %>% select(2) %>% pull()
dt_to<-ranges %>% filter(labels == input$period_corrplot) %>% select(3) %>% pull()
aux<-subset(log_ret_df, Date>=dt_from & Date <= dt_to)

corr <- round(cor(aux[-1]),12)
ggcorrplot(corr, hc.order = TRUE, type = "lower",lab = TRUE)
})

renderPlot({
dt_from<-ranges %>% filter(labels== input$period_corrplot) %>% select(2) %>% pull()
dt_to<-ranges %>% filter(labels == input$period_corrplot) %>% select(3) %>% pull()
log_ret_aux<-long_log_ret[(long_log_ret$Date>=dt_from &long_log_ret$Date<=dt_to),]

ggplot(log_ret_aux, aes(x = Date, y = value)) +
  geom_line() +
  facet_wrap(~variable, scales = "free") +
  ggtitle("Log Returns")

})
```


### Summary and Statistical Tests
```{r test_prep, echo = F}
selectInput("period_table", "Choose a period:",
      list(`Log Returns Period` = list( "GFC", "Post-GFC", "Low oil prices","Fed Rates Increase","Covid"))
    )
renderTable({
dt_from<-ranges %>% filter(labels== input$period_table) %>% select(2) %>% pull()
dt_to<-ranges %>% filter(labels == input$period_table) %>% select(3) %>% pull()

long_log_ret<-long_log_ret[(long_log_ret$Date>=dt_from &long_log_ret$Date<=dt_to),]
log_ret_df<-log_ret_df[(log_ret_df$Date>=dt_from & log_ret_df$Date<=dt_to),]  
  
summary_df<-long_log_ret %>% group_by(as.factor(variable)) %>%
  get_summary_stats(type = "common") %>%dplyr::select(-variable,n) %>% 
  rename("variable" = 1)%>%
  data.table::transpose( keep.names = "col",make.names = "variable")
rownames(summary_df)<-summary_df[,1]
summary_df<-summary_df[-1]

summary_df[c("skewness", "kurtosis"),] <- rbind(c(round(sapply(log_ret_df[c(2:6)],skewness),3),
                                                  round(sapply(log_ret_df[c(2:6)],kurtosis),3)))
colnames(summary_df) <- stylized_names

ar_1_list <- list()
q_10_list<- list()
q2_10_list<- list()
ma_1_list <- list()
arch_4_list<-list()
shapiro_norm_list<-list()
adf_list<-list()
jarque_bera_list<-list()
garch_list<-list()
garch_q_10<-list()
garch_q2_10<-list()

for(i in 2:length(log_ret_df)){
  val<-log_ret_df[,i] %>%pull( )
  lag_val<-lag(val,1)
  res<-lm(val~lag_val)$residuals
  fit_auto <- auto.arima(val, ic = c("bic"), stationary = TRUE)
  opt_ar<-ifelse(length(fit_auto$coef)>0,fit_auto$arma[1],0)
  opt_ma<-ifelse(length(fit_auto$coef)>0,fit_auto$arma[2],0)
  garch_orders<-best_garch(fit_auto)
  model_specification <- ugarchspec(mean.model = list(armaOrder = c(opt_ar,opt_ma), include.mean = TRUE),
                                        variance.model = list(garchOrder = c(garch_orders[1],garch_orders[2])))
  garch_fit_res<-residuals(ugarchfit(model_specification,val),standardize=T) %>% as.vector()
  shapiro_norm_list[[i-1]]<-shapiro.test(sample(val,size=min(5000,length(val)),replace = F))
  adf_list[[i-1]] <- tseries::adf.test(val)
  ma_1_list[[i-1]] <- opt_ar
  ar_1_list[[i-1]] <- opt_ma
  q_10_list[[i-1]]<-Box.test(fit_auto$residuals, lag = 10, type =  "Ljung-Box")
  q2_10_list[[i-1]]<-Box.test(fit_auto$residuals^2, lag = 10, type =  "Ljung-Box")
  arch_4_list[[i-1]]<- ArchTest(fit_auto$residuals, lags = 4, demean = FALSE)
  jarque_bera_list[[i-1]]<-jarque.bera.test(val)
  garch_list[[i-1]]<-garch_orders
  garch_q_10[[i-1]]<-Box.test(garch_fit_res, lag = 10, type =  "Ljung-Box")
  garch_q2_10[[i-1]]<-Box.test(garch_fit_res^2, lag = 10, type =  "Ljung-Box")
}

res_matrix<-data.frame(matrix(ncol=length(stylized_names),nrow=20))
x <- c("name", "age", "gender")
colnames(res_matrix) <- stylized_names
rownames(res_matrix)<-c("Shapiro Normality","Shapiro Normality PVal","ADF","ADF Pval","Jarque-Berra","Jarque Berra Pval","AR order","MA order","Q(10)","Q(10) Pval","Q2(10)","Q2(10) Pval","ARCH(4)","ARCH(4) Pval","GARCH Order(p)","ARCH Order (q)","Garch Q(10)","Garch Q(10) Pval","Garch Q2(10)","Garch Q2(10) Pval")

for (i in 1:length(stylized_names)){
  res_matrix[1,i]<-round(shapiro_norm_list[[i]]$statistic,2)
  res_matrix[2,i]<-shapiro_norm_list[[i]]$p.value
  res_matrix[3,i]<-round(adf_list[[i]]$statistic,2)
  res_matrix[4,i]<-adf_list[[i]]$p.value
  res_matrix[5,i]<-round(jarque_bera_list[[i]]$statistic,3)
  res_matrix[6,i]<-jarque_bera_list[[i]]$p.value
  res_matrix[7,i]<-ar_1_list[[i]]
  res_matrix[8,i]<-ma_1_list[[i]]
  res_matrix[9,i]<-round(q_10_list[[i]]$statistic,2)
  res_matrix[10,i]<-q_10_list[[i]]$p.value
  res_matrix[11,i]<-round(q2_10_list[[i]]$statistic,2)
  res_matrix[12,i]<-q2_10_list[[i]]$p.value
  res_matrix[13,i]<-round(arch_4_list[[i]]$statistic,2)
  res_matrix[14,i]<-arch_4_list[[i]]$p.value
  res_matrix[15,i]<-garch_list[[i]][1]
  res_matrix[16,i]<-garch_list[[i]][2]
  res_matrix[17,i]<-round(garch_q_10[[i]]$statistic,2)
  res_matrix[18,i]<-garch_q_10[[i]]$p.value
  res_matrix[19,i]<-round(garch_q2_10[[i]]$statistic,2)
  res_matrix[20,i]<-garch_q2_10[[i]]$p.value
}

for (i in 1:(length(res_matrix))){
  row <- as.character(res_matrix[,i])
  for (j in seq(from = 1, to = length(row)-1,by = 2)){
    if(!j %in% c(7,8,15,16)){
      coef<-round(as.numeric(row[j]),4)
      pval<-as.numeric(row[j+1])
      row[j]<-ifelse(pval<=0.001,paste(coef,"***",sep=""),
                     ifelse(pval<=0.01,paste(coef,"**",sep=""),
                            ifelse(pval<=0.05,paste(coef,"*",sep=""),as.character(coef))))
      row[j+1]<-as.character(round(pval,5))
    }
  }
  res_matrix[,i]<-row
}

final_df<-summary_df %>% rbind(res_matrix)
final_df
},options = list(scrollX = TRUE),
rownames = TRUE, width="100%",spacing="xs")
```

## Spillovers

All functions are documented in the frequencyConnectedness package on [Cran](https://cran.r-project.org/web/packages/frequencyConnectedness/frequencyConnectedness.pdf).

```{r VAR, echo = F,eval=F}
opt_lag<-VARselect(log_ret_df[,-c(1)],type="const",lag.max=15)
df_opt<-data.frame(t(opt_lag$criteria))
colnames(df_opt)<-c("AIC","HQ","SC","FPE")
rownames(df_opt)<-paste("Lag",rep(1:nrow(df_opt)))

for (i in 1:length(opt_lag$selection)){
  df_opt<-format_cells(df_opt,opt_lag$selection[i],i,"bold")
}

df_opt%>% 
  kbl()%>%
  kable_classic_2(full_width = F)
```


```{r AR for VAR, echo = F}
### Testing of AR in residuals for VAR(1)
est_lag1<-vars::VAR(log_ret_df[-1], p = 1, type = "const")
ar_lag1_list<-list()
for(i in 1:length(est_lag1$varresult)){
  res<-est_lag1$varresult[i][[1]]$residuals
  ar_lag1_list[[i]] <- arima( res, order = c(1, 0, 0))
}
```

```{r stargazer4,results='asis',echo=F,eval=F}
stargazer(ar_lag1_list,type="html",align = T)
```

```{r main_sp_tables, echo = F,eval=F}
sp_tbl<-list.files("./data/sp_data/total") 

for (i in sp_tbl){
  sp_table<-readRDS(paste0("data/sp_data/total/",i))
  print(i)
  print(sp_table)
}
```

```{r echo=F}
sp_lag1_total<-readRDS("./data/sp_data/total/sp_lag1.rds")
sp_lag1<-readRDS("./data/sp_data/sp_w100_lag1.rds")
sp_lag1_roll<-readRDS("./data/sp_data/plots/sp_plot_w100_lag1.rds")
```

### Spillover plots
#### Total
```{r total,echo=F}
print(sp_lag1_total)
plot(sp_lag1_roll)
```

#### Net

```{r net,echo=F}
net(sp_lag1_total)
plotNet(sp_lag1)
```

#### From

```{r from,echo=F}
from(sp_lag1_total)
plotFrom(sp_lag1)
```

#### To

```{r to,echo=F}
to(sp_lag1_total) 
plotTo(sp_lag1)
```


```{r pairwise, include=F}
z<-plotPairwise(sp_lag1)
colnames(z)<-c("NG-Crude","NG-Gasoline","NG-Heating",
               "NG-Diesel","Crude-Gasoline","Crude-Heating",
               "Crude-Diesel","Gasoline-Heating","Gasoline-Diesel","Heating-Diesel")
```

```{r fig.height = 8,echo=F,eval=F}
pairwise(sp_lag1_total)
plot(z,main="Pairwise spillovers")
```


## Frequency Dependent Spillover

Get the frequency connectedness on partition roughly corresponding to movements of 1-5,5-20,20-200,200-inf days.
Using window 100, lag 1
```{r frequency,echo=F}
bounds <- c(pi+0.00001, pi/5, pi/20, pi/200, 0)
sp_BK12<-spilloverBK12(est_lag1, n.ahead = 100, no.corr = F, partition = bounds)
print("Frequency Bands")
print(sp_BK12)
# print("Total (collapsed bands)")
# collapseBounds(sp_BK12, 1:4)
sp_roll_BK12<-readRDS("data/sp_data/sp_roll_freq_lag1.rds")
ovs<-overall(sp_roll_BK12)
ovs_freq<-data.frame(matrix(NA,0,3))
colnames(ovs_freq)<-c("series","date","val")
mains<-c("1-5","5-20","20-200","200-inf")
for(j in 1:length(ovs)){
  i<-ovs[[j]] %>% as.data.frame()
  ov_sp <- cbind(rep(mains[j],nrow(i)),rownames(i), i)
  colnames(ov_sp)<-c("series","date","val")
  ovs_freq<-rbind(ovs_freq,ov_sp)
}
ovs_freq$date<-as.POSIXct(ovs_freq$date)
rownames(ovs_freq) <- 1:nrow(ovs_freq)
ovs_freq$series = factor(ovs_freq$series, levels=mains)
freq_plot<-ggplot(ovs_freq,aes(x=date,y=val, colour=series)) + 
  geom_line() + 
  scale_y_continuous(limits = c(0, 80))+
  scale_x_datetime(labels = date_format("%Y"),date_breaks = "1 year")+
  theme(axis.text.x=element_text(angle=60,hjust=1))+
  theme(axis.title.x=element_blank())+
 theme(legend.position = "none")

total_lag1<-readRDS("data/sp_data/sp_w100_lag1.rds")
total_lag1<-overall(total_lag1)[[1]] %>% as.data.frame()
total_lag1 <- cbind(rownames(total_lag1), total_lag1)
colnames(total_lag1)<-c("date","val")
rownames(total_lag1) <- 1:nrow(total_lag1)
total_lag1$date<-as.POSIXct(total_lag1$date)

total_plot<-ggplot(total_lag1,aes(x=date,y=val)) + 
  geom_line() + 
  scale_y_continuous(limits = c(0, 80))+
  scale_x_datetime(labels = date_format("%Y"),date_breaks = "1 year") +
  theme(axis.text.x=element_text(angle=60,hjust=1))+
  theme(axis.title.x=element_blank())

```
### Frequency plots
Red 1-5

Green 5-20

Blue 20-200

Purple 200-inf (Was not plotted in Fig.4 BK2019)
```{r fig.height = 8, fig.width = 8}
grid.arrange(total_plot, freq_plot, nrow = 2)
```

#### Correspondiong spikes
```{r spikes}
total_lag1 %>% arrange(desc(val)) %>% head(5)
ovs_freq%>%dplyr::filter(series=="200-inf") %>% arrange(desc(val)) %>% head(5)
```


### Frequency spillovers for selected ranges
```{r period_sp_freq,echo=F,message=T,warning=T}

selectInput("period_freq", "Choose a period:",
      list(`Freq Period` = list("GFC", "Post-GFC", "Low oil prices","Fed Rates Increase","Covid"))
    )

renderPlot({
dt_from<-ranges %>% filter(labels== input$period_freq) %>% select(2) %>% pull()
dt_to<-ranges %>% filter(labels == input$period_freq) %>% select(3) %>% pull()

sp_roll_BK12<-readRDS("data/sp_data/sp_roll_freq_lag1.rds")
ovs<-overall(sp_roll_BK12)
ovs_freq<-data.frame(matrix(NA,0,3))
colnames(ovs_freq)<-c("series","date","val")
mains<-c("1-5","5-20","20-200","200-inf")
for(j in 1:length(ovs)){
  i<-ovs[[j]] %>% as.data.frame()
  ov_sp <- cbind(rep(mains[j],nrow(i)),rownames(i), i)
  colnames(ov_sp)<-c("series","date","val")
  ovs_freq<-rbind(ovs_freq,ov_sp)
}
ovs_freq$date<-as.POSIXct(ovs_freq$date)
rownames(ovs_freq) <- 1:nrow(ovs_freq)
ovs_freq$series = factor(ovs_freq$series, levels=mains)
ovs_freq<-ovs_freq %>% filter(date>=dt_from,date<=dt_to)
freq_plot<-ggplot(ovs_freq,aes(x=date,y=val, colour=series)) + 
  geom_line() + 
  scale_y_continuous(limits = c(0, 80))+
  scale_x_datetime(labels = date_format("%Y-%m"),date_breaks = "1 month")+
  theme(axis.text.x=element_text(angle=60,hjust=1))+
  theme(axis.title.x=element_blank())
freq_plot
})

```

## PCA
```{r pca_stats,echo=F}
print("Correlation of normalized indexes")
ggcorrplot(pca_corr_norm,lab = TRUE)
print("Correlation of returns")
ggcorrplot(pca_corr_ret, lab = TRUE)
head(pca_vals) %>% 
    kbl()%>%
    kable_classic_2(full_width = F)
```








